---
title: 'NaiveBayes: Analysis of Race on COVID-19 Conditional Death Probabilities'
author: "Collin Fabian"
date: "3/21/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('/Users/fabian_coll/Desktop/PyProjects/COVID-Race-Income-Analysis/naiveBayes')
```

We must first load in our data
```{r}
library(e1071)
patient_df <- read.csv('patient_cases_by_race.csv')
```

```{r}
patient_df$hosp_yn <- factor(patient_df$hosp_yn)
head(patient_df, n = 10)
```

```{r}
train.index <- sample(c(1:dim(patient_df)[1]), dim(patient_df)[1]*0.6)  
train.df <- patient_df[train.index, ]
valid.df <- patient_df[-train.index, ]
```

Now, we can use the Naive Bayes Classifier for Discrete Predictors to calculate the probability of death based on race.
```{r}
race.nb <- naiveBayes(Race ~ death_yn, data = train.df)
race.nb$tables$death_yn
```
As we can see, the Races ranked in order from highest percentage of death to lowest is:\
  1. AmericanIndian/NativeIslander\
  2. Black\
  3. White\
  4. Asian\
  5. Other\
  6. NativeHawaiian/PacificIslander\
  7. Hispanic\
 \
 
```{r}

```
Now, lets use Naive Bayes again to get P(Hosp_yn | Race & death_yn = Yes). This will tell us the probability that a patient who died had access to a hospital given their Race.
```{r}
race.nb2 <- naiveBayes(Race ~ hosp_yn, data = subset(train.df, death_yn=='Yes'))
race.nb2$tables$hosp_yn
```
From our results, out of all the patients who died from COVID-19, the probabilities of getting treatment at a hospital based on Race are listed in order from lowest to highest:\
  1. White\
  2. Other\
  3. AmericanIndian/NativeIslander\
  4. NativeHawaiian/PacificIslander\
  5. Hispanic\
  6. Black\
  7. Asian\
  
  





